\boldmath Supersymmetric Dirac-Born-Infeld Axionic Inflation and Non-Gaussianity

### 2018-10-29 SUSY_DBI_referee_reply.tex

\def\co{coannihilation~}
\def\G{\tilde G}
\def\mP{M_{\rm Pl}}

\begin{document}

 \begin{center}
 Reply to report of referee
  \end{center}

{\it We thank the referee for his/her comments and suggestions.
In the revised version we have taken account of these comments and suggestions.
Below are our brief replies to the referee report.}

@@

\noindent
Referee:\\
 The paper studies inflation and non-Gaussianity in a two-field supersymmetric DBI model. The paper is of some interest, but I have several questions and a few suggestions for improvement before I can formulate a recommendation.\\

1) What value of m in (3.11) is taken in the concrete numerical evaluations?\\
 \noindent {\it Reply\\
 In the analysis we have taken $m=6$. We also added a corresponding note to the paper in Sec.(6). }

2) Nearly all relations in the paper are given in an implicit way, (for example keeping derivatives of the superpotential unspecified). In my opinion, the paper would improve if it showed the following explicit results:\\

 {\it Reply\\
 The derivatives of the superpotential first appear in Eq. (3.12) because their explicit forms were not needed in the
 immediate discussion that follows Eq. (3.12). They were made explicit in Eqs.(3.20), (3.21), (3.23)-(3.26). 
 However, we are happy to make them explicit immediately after Eq. (3.12) which we have done in the two unnumbered equations 
 after Eq.(3.12) in the revised version of the paper.
 Similarly we have added a line after Eq.(3.5) directing the reader to section 8 (an appendix) where the covariant derivatives
 are defined.
 }

@@

2.2) A similar analysis should be done for the combination  $a_+$. After eq. (3.26), the paper states "$a_+$ undergoes fast roll.... thus we suppress $a_+$". This point should be clarified. My understanding is that this sentence means that $a_+$ has a much larger mass than a?, so we can consistently set $a_+$ = 0. This should be shown explicitly, by computing explicitly the mass of this field (in terms of the parameters in the model) and showing that it is indeed greater than H.\\

@@

\noindent {\it Reply\\
This is one of the central points of the analysis, i.e., decomposition of the potential into slow roll vs fast roll where the slow roll potential
 depends  only on $a_-$ and the remaining part of the potential is controlled by $a_+$ and is fast roll. The reason we did not elaborate
 too much on  it in this paper is because this topic was  dealt with at great length in our previous paper  ``Evidence for Inflation in an Axion Landscape,''  JHEP {\bf 1803}, 121 (2018). In this work it was shown that $a_+$ has a much larger mass than $a_-$ and further
 that the $a_+$ rolls an order of magnitude or more  faster than $a_-$ as shown by blue lines (fast roll) vs red lines (slow roll) in Fig.1
 and Fig.2 of JHEP {\bf 1803}, 121 (2018). However, we have added an additional text in the revised version of the paper 
 to explain this further. This is done at the end of the paragraph after Eq. (3.27).
}

2.3) It would be useful to see the explicit form of $V (a_-)$, when the other fields are set to the minimum. In this way the reader can see the explicit form of the inflaton potential used in the numerical evolutions. If many terms are present, it would be useful to single out the dominant ones (I would expect the usual cosine potential, plus corrections; please provide the expression for the dominant term and the leading correction).
2.4) The parametrization (3.19) suggests that $\rho_k = a_k = 0$ in the minimum of the potential. Therefore, as it is standard, the axion scale 
$f_k$ is obtained from the minimization of the scalar potential. Please provide the explicit relation that gives $f_k$ in terms of the parameters of the model.\\
\noindent {\it Reply\\
This is an interesting question.  Indeed for conventional inflation models an explicit form of the potential plays a central role in the 
analysis. However, the potential by itself is not the whole story in the analysis for the Dirac-Born-Infeld case. Here the entire
Lagrangian enters in the analysis and the potential in isolation plays no role in the analysis. This can be seen from the Friedman equations
Eqs.(4.9) and (4.10) of the paper 
where the full Lagrangian and not  just the potential enters. Only in the case when we assume canonical kinetic energy and
no dependence of the potential on the derivatives of the fields that one recovers the conventional Friedman equations where 
the potential appears as shown in Eqs. (4.12).  However, we thank the referee for bringing this point up and we have added 
a few lines to explain this point after Eq. (4.12). Regarding $f_k$, 
 a closed form solution for $f_k$ cannot be exhibited since it satisfies
a polynomial constraint of high order. However,  we have exhibited the explicit form of 
$f$ on the input parameters in Eq. (3.22) in a perturbation expansion on the symmetry breaking part.
}

@@

3) There are different shapes of non-gaussianity (the equilateral and the local shapes being the most studied ones), and the limits on non-gaussianity strongly depend on the shape. By shape, we mean the dependence of the bispectrum on the ratios $k_2/k_1$ and
$k_3/k_1$. The paper cites [64,65,66]. Eq. 11 of [64], which is also eq. 1 of [65], and eq. 17 of [66] defines the so called local shape. This is different from the shape of non-Gaussianity obtained in DBI models, which is very close to the equilateral shape (see for instance Section 2.1 of the Planck paper 1502.01592). Moreover the works [64,65,66], although very important, are rather outdated by now. So, this part of the paper should be updated, and the value of $f_{NL}$ should be compared to the current limits (see 1502.01592) or future forecasts on equilateral non-Gaussianity, not the local one.

\noindent {\it Reply\\ 
We agree with the referee
that our references to the analysis of non-Gaussianity from the Planck experiment need updating. We have
done so in the revised version.  The changes made appear after Eq.(5.12) and near the end of section 6. We have included
the reference to 1502.01592 and specified the current limits on $f_{NL}^{equil}$.
}

4) I don't understand eq.(6.1). The conventional way to find the end
of inflation is to find the moment at which $\ddot a = 0$ (second derivative of the
scale factor with respect to time). The number of e-folds during inflation is
typically defined from $a(N ) \equiv a_{end} e^{-N}$ , which gives  $\frac{dN}{dt} = -H$ (where $a_{end}$ 
is the value of the scale factor at the end of inflation, and H is the Hubble
rate). In this way, $N = 0$ at the end of inflation, while N is positive at
early times during inflation. How is eq. (6.1) used to determine the end of
inflation ?  What is the purpose of setting $\theta_n = \frac{1}{16}$?  Why don't the authors 
find from their numerical evolutions the point at which $\ddot a = 0$ ?\\
\noindent {\it Reply\\ 
We agree with the referee  regarding the conventional way to find  the end of inflation.
For the Dirac-Born-Infeld the numerical integration is time intensive especially for the very large scan
involving $0.5\times 10^6$ parameter sets that we carried out.  For this reason we approximated the end of inflation the way we did
which is much faster to implement than the $\ddot {a}=0$ constraint. However, we have taken time to 
carry out the analysis using the $\ddot{a}=0$ constraints and our analysis shows that the results of our approximate
procedure and the conventional ones agree for 97\% of the parameter points. In any case we have replaced the old 
Figs (1) and (2) with new ones using $\ddot {a}=0$ constraint and a comparison of the new and old figs shows that
they are remarkably similar. Further, we have replaced Eq.(6.1) with the new constraint.
}

5) The scan shown in Figure 1 contains no real information, unless one
specifies what is varied among the different points. The only information I
could find on this regard is "we allow $\alpha_1$ to vary" at the end of page 12.
How is it varied ? (in which range ? Is it a uniform sampling in linear units
? In log units ?) What is chosen for the other parameters of the model
? (?,?,A,T). How is the correct normalization of the scalar perturbations
enforced? (this is typically done by fixing the overall scale of the scalar
potential). Are subleading terms in $\frac{1}{T}$  important ? The scan strategy should T
be completely described, so that any reader, if interested, could precisely reproduce those figures.\\

@@

\noindent {\it Reply\\
We agree that the distribution used to generate the dataset from Figs.(1, 2) is important to be
mentioned in the paper. We added a complete specification to the caption of Fig.(1).
}

6) The paper motivates the model from the trans-Planckian problem of axion inflation (the fact that we must take the axion scale 
$f > M_p$ in the original potential of natural inflation). It would be very interesting to provide also a scan of the values of f obtained for the various realizations presented in figures 1 and 2. Does this scan extend in the sub-Planckian region ? (the scan should be easily done, once the above questions 2.4 and 5 are answered).\\

\noindent {\it Reply\\
We agree with the suggestion of the referee. We have added a new figure 4 and a short discussion of it at the end of section 6 where we have presented a scatter plot of $f / M_\text{P}$ vs $n_s$ and $f / M_\text{P}$ vs $r$ for the data sets of figures 1 and 2. Here we see that $f$
 lies in the sub-Planckian region for a signifiant set of parameter points.}

### 2018-10-29 SUSY_DBI.tex
New end of inflation (#14)
* Closes #8 

I did some thinking related to the two options you suggested.

In this regard since we have the analysis now using \ddot a=0 condition, we can replace figs 1 and 2. Fig. 3 we leave as is because it is just an illustrative figure.

So I have worked on completing the reply to the referee report and the new version is attached. The text in red in item 4) is new and so please read and comment.

I have also added some text in the revised version of the paper related to item 4 and item 6.

   because $|W_{sb}| << |W_s|$. In the analysis of this there is just one pair of axion field so we have only $a_+$ and $a_-$ 

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Pressure, density and inflation equations \label{sec4}}
  The pressure $p$ and density $\rho$ are defined in terms of the stress tensor by
  \begin{equation}\label{DisplayFormulaNumbered:eq.pressure}
    p= \frac{1}{3} \sum_{i=1}^3 T^{ii}\,,
    ~~\rho =T^{00},
  \end{equation}
  where $T^{\mu\nu}$ $(\mu=0, 1, 2, 3)$ is the stress tensor and is given by
  \begin{equation}
    T^{\mu \nu}=g^{\mu \nu} \mathcal{L}\left(\phi_k,\partial_\mu \phi_k \partial^\mu \phi_k \right)-2\frac{\delta \mathcal{L}\left(\phi_k,\partial_\mu \phi_k \partial^\mu \phi_k \right)}{\delta \left(\partial_\mu \phi_k \partial^\mu \phi_k \right)}\partial^\mu \phi_k \partial^\nu \phi_k,
    \label{tmunu}
  \end{equation}
  and we use the metric $\eta^{\mu\nu}= {\rm diag}(-1, 1, 1, 1)$.
  In the analysis we assume space to be homogeneous and isotropic, so that $\partial_i\phi_k = 0$ for $1 \leq i \leq 3$ and 
	$\partial_0 \phi_{k} = \dot{\phi}_k$,
  one finds that Eqs. (\ref{DisplayFormulaNumbered:eq.pressure}) and (\ref{tmunu})
  for pressure and density become:
  \begin{equation}\label{DisplayFormulaNumbered:eq.pressure.cosmology}
    p = \mathcal{L}\left(\phi_k, - \beta_k \right),
  \end{equation}
  and
  \begin{equation}\label{DisplayFormulaNumbered:eq.density.cosmology}
    \rho = -\mathcal{L}\left(\phi_k, - \beta_k\right) + 2\sum_{k}\frac{\delta \mathcal{L}\left(\phi_k, - \beta_k \right)}{\delta \beta_k}
    \beta_k,
  \end{equation}
  where $\beta_k = \dot \phi_k^2$.
  These relations are valid for non-canonical kinetic terms.
  The Friedman equations are given by
  \begin{align}
    \dot \rho = -3H (p + \rho),
    \label{fried1}
  \end{align}
  \begin{align}
    3 M_\text{P}^2 \frac{\dot R^2}{R^2} = \rho,
    \label{fried2}
  \end{align}
  \begin{align}
    \dot\rho= - 6H \sum_{k} \frac{\partial L}{\partial \beta_k} \beta_k\,.
    \label{rhodot1}
  \end{align}
  Further taking the time derivative of Eq.~(\ref{DisplayFormulaNumbered:eq.density.cosmology})
  which along with Eq.~(\ref{rhodot1}) gives
  \begin{align}
    &\sum_{k}\Big[ \left(2 \frac{\partial^2 L}{\partial \beta_k \partial \beta_k} \beta_k + \frac{\partial L}{\partial \beta_k}
    \right)\dot \beta_k
    - \frac{\partial L}{\partial \phi_k} \dot \phi_k + 6H \frac{\partial L}{\partial \beta_k} \beta_k
    + 2 \frac{\partial^2 L}{\partial \phi_k\partial \beta_k} \beta\dot \phi_k \Big]=0.
    \label{motion}
  \end{align}
  Next we focus on the slow-roll part where we keep only the field $a_-$. In this case Eqs. (\ref{fried2}) and (\ref{motion}) can be written in
  the following form
  \begin{align}
    3 M_\text{P}^2 \frac{\dot R^2}{R^2} = 2 \dot a_{-}^2 \frac{\partial L}{\partial \dot a_{-}^2} - L\,,
    \label{inf-1a}
  \end{align}
  \begin{align}
    2 \left[2 \frac{\partial^2 L}{\partial \dot a_{-}^2 \partial \dot a_{-}^2} \dot a_{-}^2 + \frac{\partial L}{\partial \dot a_{-}^2}
    \right] \ddot a_{-} -
    \frac{\partial L}{\partial a_{-}} + 6H \dot a_{-} \frac{\partial L}{\partial \dot a_{-}^2}
    + 2 \frac{\partial^2 L}{\partial a_{-} \partial \dot a_{-}^2} \dot a_{-}^2
    =0
    \label{inf-2a}\,.
  \end{align}
  For the case of canonical kinetic energy and no dependence of the potential on time derivative of field, i.e.,
  \begin{align}
    L= -\frac{1}{2} \partial_\mu a_{-}\partial a_{-}^{\mu} - V(a_-)
  \end{align}
  Eqs. (\ref{inf-1a}) and  (\ref{inf-2a}) 
   reduce to the following
  \begin{align}
    3 M_\text{P}^2 \frac{\dot R^2}{R^2} = \frac{1}{2} \dot a_{-}^2 + V(a_-)\,,\non
    \ddot a_{-} + 3 H \dot a_{-} + V'(a_{-})=0\,,
  \end{align}
  which are correctly the relation for slow roll for the case when one has canonical kinetic energy and the potential
  is velocity independent. However, in our case we have more terms.

@@

  Non-Gaussianity is defined by the three-point correlation function of perturbations involving
  three scalars, two scalars and a graviton, two gravitons and a scalar and three gravitons~\cite{Maldacena:2002vr}.
  Since the work of~\cite{Maldacena:2002vr}  there has been
  a significant number of further analyses (see, e.g., \cite{Acquaviva:2002ud,Creminelli:2003iq,
  Silverstein:2003hf,Gruzinov:2004jx,Alishahiha:2004eh,Chen:2005fe,Creminelli:2005hu,
  Seery:2005wm,Babich:2004gb,Lyth:2005fi,Chen:2006nt, Huang:2007hh,Langlois:2008wt}).
  The dominant non-Gaussianity
  arises from the correlation function of three scalar perturbations. Thus for scalar perturbation $\zeta(\vec k)$ non-Gaussianity
  is defined by
  \begin{align}
    \left< \zeta(\vec k_1) \zeta(\vec k_2) \zeta(\vec k_3) \right> = (2\pi)^7 \delta^3(\vec k_1+ \vec k_2+ \vec k_3) \frac{\sum_i k_i^3}{\prod_i k_i^3}
    \left[ -\frac{3}{10} f_{NL} (P_k^{\zeta})^2 \right],
  \end{align}
  where $P_k^{\zeta}$ is the scalar power spectrum and $f_{NL}$ is a measure of non-Gaussianity. 
  Non-Gaussianity depends strongly on  shape and various types of shapes have been discussed in the 
literature which include local, equilibrium and orthogonal ~\cite{Ade:2015ava} (For earlier analyses see 
\cite{Komatsu:2001rj,Komatsu:2003fd,Verde:1999ij}).
 The non-Gaussianity  we will discuss here is the one which has the shape of an equilateral triangle, 
  when $k_1=k_2=k_3$ and  is given by~\cite{Chen:2006nt}
  \begin{align}
   f^{equil}_{NL}= \frac{35}{108}\left(\frac{1}{c_s^2}-1\right) - \frac{5}{81} \left[ \left(\frac{1}{c_s^2} -1- 2\frac{z_2}{z_1}\right) + (3-2 {c}_1)z \frac{z_2}{z_1}\right] + {\cal O}(\epsilon).
    \label{equil}
  \end{align}
  where
  \begin{align}
    z_1= \beta L_{,\beta} + 2 \beta^2 L_{\beta\beta}\,,~~
    z_2= \beta^2 L_{,\beta \beta} + \frac{2}{3} \beta^3 L_{,\beta\beta\beta}\,,~~
    z= \frac{\dot z_2}{z_2 H}\,.
  \end{align}
 where the superscript ``equal''  on $f_{NL}$ indicates that we are specifically considering the shape to be of an equilateral 
 triangle.

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig4.pdf}
    \caption{Left panel: A plot of axion decay constant $f$ vs scalar spectral index $n_s$ for the
      same data set as in Fig.~(\ref{fig1}).
          Right panel: The same as the left panel except $f$ is plotted vs tensor to scalar ratio
      $r$.
          Note that $20\%$ of experimentally consistent (green) points in the figure satisfy
      $f<M_\text{P}$. Also note that we vary $f$ explicitly in this paper, which implicitly sets
      the value of $\lambda$, as can be seen from Eq.~(\ref{eq.f.series}).}
    \label{fig4}
  \end{figure}

@@

Numerical simulations performed consist of two parts: a large $0.5 \times 10^6$ points simulation in which $\alpha_2 = \alpha_3 = 0$, and a small simulation with $\alpha_2$ and $\alpha_3$ perturbed around an experimentally consistent point with the largest non-Gaussianity from the first simulation.
  The evolution of the inflaton field $a_-$ in the large simulation of Figs.~(\ref{fig1}), (\ref{fig2}), and (\ref{fig4}) is obtained by taking the effective axion Lagrangian Eqs.~(\ref{eq.axionLagrangian.I}) and (\ref{eq.axionLagrangian.II}), deriving Euler-Lagrange equation and Friedmann Eq.~(\ref{fried2}) using computer algebra, and solving them numerically. The integration is stopped when the ratio of the time derivative ($\dot n\left(t\right)$) of the e-foldings count to its time-average ($n\left(t\right) / t$) passes a threshold. Specifically,
  \begin{equation} 
  	\frac{\dot n \left(t\right)}{n\left(t\right) / t} < \theta_n,
  \end{equation} 
  where the threshold $\theta_n$ is chosen to be much less than one, and specifically
   in the analysis here it is chosen as $\theta_n = 1 / 16$. Increasing this parameter might stop integration prematurely, whereas decreasing it will worsen performance.
  An analytic expression similar to Eq.~(\ref{eq.axionLagrangian.II}), however,  cannot be derived  for  non-zero $\alpha_2$ and $\alpha_3$. In this case,  equations for the auxiliary fields $F_k$ are  solved numerically,  and the Lagrangian in terms of $\phi$ (Eq.~(\ref{Lag.23})) is
  also evaluated numerically for each set of values for ($a_-$, $\dot a_-$).  
   This process is significantly slower, therefore, in this case the simulations are only done for $29$ sets of parameter values as can be seen in Fig.~(\ref{fig3}). 
    The time of horizon exit is then found by counting $N_\text{pivot}$ e-foldings back from the end of integration, where $N_\text{pivot}$ is varied between $50$ and $60$. Note, the time of horizon exit depends on $\theta_n$, however, we find that due to $\dot n$ being small near the end of integration, the dependence is weak, and it affects the results less than the experimental uncertainty of $N_\text{pivot} \in \left[50, 60\right]$.{}

@@

@@

  Finally, Fig.~(\ref{fig4}) demonstrates that it is indeed possible to produce experimentally
  consistent inflation with $f < M_\text{P}$ in our model.

>>>>

   because $|W_{sb}| << |W_s|$. In the analysis of this paper there is just one pair of axion field so we have only $a_+$ and $a_-$ 

@@

\section{Pressure, density and inflation equations \label{sec4}}
  The pressure $p$ and density $\rho$ are defined in terms of the stress tensor by
  \begin{equation}\label{DisplayFormulaNumbered:eq.pressure}
    p= \frac{1}{3} \sum_{i=1}^3 T^{ii}\,,
    ~~\rho =T^{00},
  where $T^{\mu\nu}$ $(\mu=0, 1, 2, 3)$ is the stress tensor and is given by
    T^{\mu \nu}=g^{\mu \nu} \mathcal{L}\left(\phi_k,\partial_\mu \phi_k \partial^\mu \phi_k \right)-2\frac{\delta \mathcal{L}\left(\phi_k,\partial_\mu \phi_k \partial^\mu \phi_k \right)}{\delta \left(\partial_\mu \phi_k \partial^\mu \phi_k \right)}\partial^\mu \phi_k \partial^\nu \phi_k,
    \label{tmunu}
  and we use the metric $\eta^{\mu\nu}= {\rm diag}(-1, 1, 1, 1)$.
  In the analysis we assume space to be homogeneous and isotropic, so that $\partial_i\phi_k = 0$ for $1 \leq i \leq 3$ and 
	$\partial_0 \phi_{k} = \dot{\phi}_k$,
  one finds that Eqs. (\ref{DisplayFormulaNumbered:eq.pressure}) and (\ref{tmunu})
  for pressure and density become:
  \begin{equation}\label{DisplayFormulaNumbered:eq.pressure.cosmology}
    p = \mathcal{L}\left(\phi_k, - \beta_k \right),
  \begin{equation}\label{DisplayFormulaNumbered:eq.density.cosmology}
    \rho = -\mathcal{L}\left(\phi_k, - \beta_k\right) + 2\sum_{k}\frac{\delta \mathcal{L}\left(\phi_k, - \beta_k \right)}{\delta \beta_k}
    \beta_k,
  where $\beta_k = \dot \phi_k^2$.
  These relations are valid for non-canonical kinetic terms.
  The Friedman equations are given by
    \dot \rho = -3H (p + \rho),
    \label{fried1}
    3 M_\text{P}^2 \frac{\dot R^2}{R^2} = \rho,
    \label{fried2}
    \dot\rho= - 6H \sum_{k} \frac{\partial L}{\partial \beta_k} \beta_k\,.
    \label{rhodot1}
  Further taking the time derivative of Eq.~(\ref{DisplayFormulaNumbered:eq.density.cosmology})
  which along with Eq.~(\ref{rhodot1}) gives
    &\sum_{k}\Big[ \left(2 \frac{\partial^2 L}{\partial \beta_k \partial \beta_k} \beta_k + \frac{\partial L}{\partial \beta_k}
    \right)\dot \beta_k
    - \frac{\partial L}{\partial \phi_k} \dot \phi_k + 6H \frac{\partial L}{\partial \beta_k} \beta_k
    + 2 \frac{\partial^2 L}{\partial \phi_k\partial \beta_k} \beta\dot \phi_k \Big]=0.
    \label{motion}
  Next we focus on the slow-roll part where we keep only the field $a_-$. In this case Eqs. (\ref{fried2}) and (\ref{motion}) can be written in
  the following form
    3 M_\text{P}^2 \frac{\dot R^2}{R^2} = 2 \dot a_{-}^2 \frac{\partial L}{\partial \dot a_{-}^2} - L\,,
    \label{inf-1a}
    2 \left[2 \frac{\partial^2 L}{\partial \dot a_{-}^2 \partial \dot a_{-}^2} \dot a_{-}^2 + \frac{\partial L}{\partial \dot a_{-}^2}
    \right] \ddot a_{-} -
    \frac{\partial L}{\partial a_{-}} + 6H \dot a_{-} \frac{\partial L}{\partial \dot a_{-}^2}
    + 2 \frac{\partial^2 L}{\partial a_{-} \partial \dot a_{-}^2} \dot a_{-}^2
    =0
    \label{inf-2a}\,.
  For the case of canonical kinetic energy and no dependence of the potential on time derivative of field, i.e.,
    L= -\frac{1}{2} \partial_\mu a_{-}\partial a_{-}^{\mu} - V(a_-)
  Eqs. (\ref{inf-1a}) and  (\ref{inf-2a}) 
   reduce to the following
  \begin{align} \label{inf-canonical}
    3 M_\text{P}^2 \frac{\dot R^2}{R^2} = \frac{1}{2} \dot a_{-}^2 + V(a_-)\,,\non
    \ddot a_{-} + 3 H \dot a_{-} + V'(a_{-})=0\,,
  which are correctly the relation for slow roll for the case when one has canonical kinetic energy and the potential
  is velocity independent. However, in our case we have more terms.
  Here we emphasize that while for
  conventional inflation models the potential plays a central role in the analysis, this is not the case for the Dirac-Born-Infeld case.
  Here the potential by itself is not sufficient and the entire Lagrangian enters in the analysis.
  The potential in isolation plays no role. This can be seen from the Friedman equations
  Eqs.~(\ref{inf-1a}) and (\ref{inf-2a}) above where the full Lagrangian and not just the potential
  enters. Only in the case when we assume canonical kinetic energy and no dependence of the
  potential on the derivatives of the fields that one recovers the conventional Friedman equations
  where the potential appears as shown in Eqs.~(\ref{inf-canonical}).

@@

  Non-Gaussianity is defined by the three-point correlation function of perturbations involving
  three scalars, two scalars and a graviton, two gravitons and a scalar and three gravitons~\cite{Maldacena:2002vr}.
  Since the work of~\cite{Maldacena:2002vr}  there has been
  a significant number of further analyses (see, e.g., \cite{Acquaviva:2002ud,Creminelli:2003iq,
  Silverstein:2003hf,Gruzinov:2004jx,Alishahiha:2004eh,Chen:2005fe,Creminelli:2005hu,
  Seery:2005wm,Babich:2004gb,Lyth:2005fi,Chen:2006nt, Huang:2007hh,Langlois:2008wt}).
  The dominant non-Gaussianity
  arises from the correlation function of three scalar perturbations. Thus for scalar perturbation $\zeta(\vec k)$ non-Gaussianity
  is defined by
  \begin{align}
    \left< \zeta(\vec k_1) \zeta(\vec k_2) \zeta(\vec k_3) \right> = (2\pi)^7 \delta^3(\vec k_1+ \vec k_2+ \vec k_3) \frac{\sum_i k_i^3}{\prod_i k_i^3}
    \left[ -\frac{3}{10} f_{NL} (P_k^{\zeta})^2 \right],
  \end{align}
  where $P_k^{\zeta}$ is the scalar power spectrum and $f_{NL}$ is a measure of non-Gaussianity.
  Non-Gaussianity depends strongly on  shape and various types of shapes have been discussed in the
literature which include local, equilibrium and orthogonal ~\cite{Ade:2015ava} (For earlier analyses see 
\cite{Komatsu:2001rj,Komatsu:2003fd,Verde:1999ij}).
 The non-Gaussianity  we will discuss here is the one which has the shape of an equilateral triangle, 
  when $k_1=k_2=k_3$ and  is given by~\cite{Chen:2006nt}
  \begin{align}
   f^{equil}_{NL}= \frac{35}{108}\left(\frac{1}{c_s^2}-1\right) - \frac{5}{81} \left[ \left(\frac{1}{c_s^2} -1- 2\frac{z_2}{z_1}\right) + (3-2 {c}_1)z \frac{z_2}{z_1}\right] + {\cal O}(\epsilon).
    \label{equil}
  \end{align}
  where
  \begin{align}
    z_1= \beta L_{,\beta} + 2 \beta^2 L_{\beta\beta}\,,~~
    z_2= \beta^2 L_{,\beta \beta} + \frac{2}{3} \beta^3 L_{,\beta\beta\beta}\,,~~
    z= \frac{\dot z_2}{z_2 H}\,.
  \end{align}
 where the superscript ``equil''  on $f_{NL}$ indicates that we are specifically considering the shape to be of an equilateral
 triangle.

@@

@@

Numerical simulations performed consist of two parts: a large $0.5 \times 10^6$ points simulation in which $\alpha_2 = \alpha_3 = 0$, and a small simulation with $\alpha_2$ and $\alpha_3$ perturbed around an experimentally consistent point with the largest non-Gaussianity from the first simulation.
  The evolution of the inflaton field $a_-$ in the large simulation of Figs.~(\ref{fig1}), (\ref{fig2}), and (\ref{fig4}) is obtained by taking the effective axion Lagrangian Eqs.~(\ref{eq.axionLagrangian.I}) and (\ref{eq.axionLagrangian.II}), deriving Euler-Lagrange equation and Friedmann Eq.~(\ref{fried2}) using computer algebra, and solving them numerically. The integration is stopped when $\ddot{a} = 0$ where $a$ is the scale factor
  and one may write the end of inflation constraint as

  \begin{equation}
    \ddot{a} = a H^2(1 + \frac{\dot H}{H^2}) = 0.
  \end{equation}
  An analytic expression similar to Eq.~(\ref{eq.axionLagrangian.II}), however,  cannot be derived  for  non-zero $\alpha_2$ and $\alpha_3$. In this case,  equations for the auxiliary fields $F_k$ are  solved numerically,  and the Lagrangian in terms of $\phi$ (Eq.~(\ref{Lag.23})) is
  also evaluated numerically for each set of values for ($a_-$, $\dot a_-$).  
   This process is significantly slower, therefore, in this case the simulations are only done for $29$ sets of parameter values as can be seen in Fig.~(\ref{fig3}). 
    The time of horizon exit is then found by counting $N_\text{pivot}$ e-foldings back from the end of integration, where $N_\text{pivot}$ is varied between $50$ and $60$.

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig4.pdf}
    \caption{
      Left panel: A plot of $f / M_{\rm P}$ as a function of $n_s$ which is in the allowed range of
      experiment and $N_{\rm pivot}$ in the range $50-60$ for the same data set as in
      Fig.~(\ref{fig1}).
      Right panel: The same as the left panel except that the plot is of $f / M_{\rm P}$ as a
      function of $r$.
      We note that the plots show that there are a large number of parameter points where
      $f / M_{\rm P} < 1$ consistent with all the experimental data.}
    \label{fig4}
  \end{figure}

@@

  Finally in Fig.(\ref{fig4}) we exhibit that the analysis contains a significant part of the parameter space where
  inflation consistent with the desired number of e-foldings in the range $[50,60]$ and $n_s$ and $r$ consistent
  with Planck data can be gotten in the axion inflation model with $f/M_{\rm P} <1$. Thus the left panel
  of Fig. (\ref{fig4}) exhibits $f/M_{\rm P}$ as a function of $n_s$ in the range allowed by the Planck data constraint of Eq.(\ref{data})
  and the right panel of Fig. (\ref{fig4}) gives $f/M_{\rm P}$ as a function of $r$
  containing the range in $r$ consistent with the Planck data constraint of Eq.(\ref{data}).
